Lesson 3
========================================================
# setwd('/Users/Anaya/repos/udacity_data_analyst/project_4/eda-r-scripts/lesson3)
***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
#read.delim('pseudo_facebook.tsv' )
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
names(pf) # pass variables 
#qplot( x = factor(dob_day), data = pf, bins=30) + scale_x_continuous(breaks=1:31) #histogram
#ggplot(aes(x = dob_day), data = pf) + 
#  geom_histogram(binwidth = 1) + 
#  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
It looks like the dob of the 1st is large, this could be due to the fact that nobody wants' to post there birthday and gives an inaccurate representation 
of what day their own dob is. 
***

### Moira's Investigation
Notes: see research paper. 
***

### Estimating Your Audience Size
Notes: I shared a post regarding a simspons clip and foudn that i had overwhelmingly more feed back as more likes came through the post
It seems that facebook starts sharing the posts to people who are actively looking at my FB feed first and if they like my post 
then it begins to spread to people who moderately look at my feed and so on and so forth. I'm pretty sure that over 300 peopel saw this post because i received 
higher then normal likes pertaining the post. Since i have 700 friends and 300 people saw my post I would say that 300/700 saw this or 43%

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:
 people underestimate the size of their audience. Plottihng hisograms. 
***
### Faceting
Notes:

```{r Faceting}

qplot(x = factor(dob_day), data = pf) + 
  scale_x_discrete(breaks = 1:31) + 
  facet_wrap(~dob_month, ncol = 3) #this takes one hisogram and splits into tweleve months
# factet_grid(formula) takes vertical ~ horizontal directions passing two or more variables 
```

#### Let’s take another look at our plot. What stands out to you here?
Response:
This histogram plot looks pretty uniform to me except the first month. Also there seems to be five months
that don't include day 31 because there aren't 31 days in this 
***

### Be Skeptical - Outliers and Anomalies
Notes: and outlier might be accurate about extreme case. Sometimes outliers represent bad data
or limitations of your data, or, in other cases extremed might be replaced with normal value. 


***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response: Probably bad data about an extreme case. 

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count, data= pf, xlim =c(0,1000)) #xlim gives me the start and end pos i want. 

#another way to do this 
qplot(x=friend_count, data = pf) + scale_x_continuous(limits = c(0,1000))
```

#### How is this plot similar to Moira's first plot?
Response: It's skewed like Moira's

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x=friend_count, data= pf, xlim =c(0,1000)) #xlim gives me the start and end pos i want. 

#another way to do this 
qplot(x=friend_count, data = pf) + scale_x_continuous(limits = c(0,1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))

# seperates the counts of data into 0 or 1 
qplot(x = friend_count, data = pf, binwidth = 10) + 
  scale_x_continuous(limits = c(0,1000), 
                     breaks = seq(0,1000,50)) + 
  facet_wrap(~gender, ncol =2) 
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
#we need to get rid of NA 
qplot(x= friend_count, data = subset(pf, !is.na(gender)),
      bindwidth = 10) + 
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50)) + 
  facet_wrap(~gender, ncol = 2)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender) # table of variables 
by(pf$friend_count, pf$gender, summary) #by function creates a summary of data
```

#### Who on average has more friends: men or women?
Response: female

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response:Since more users have less then 100 friends it would seem natural that the median would be a better metric to use for the overall amount of males and friends

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, bindwidth = 30, color = I('black'), fill = I('#099DD9')) 

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, binwidth = 1, color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(breaks = 1:7)
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, 
      xlab = 'Number of years using Facebook', 
      ylab = 'Number of users in sample', 
      color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(breaks = seq(1,7,1), lim = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot( x= age, data = pf, bindwidt = 1,
       xlab = "Age of Facebook Users",
       ylab = "Number of Age of Users",
       color = I('black'), 
       fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(0,120,10))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
install.packages('gridExtra')
library(gridExtra)
qplot(x = friend_count, data = pf)

summary(pf$friend_count)
summary(log10(pf$friend_count + 1)) # base 10 
summary(sqrt(pf$friend_count)) #sqrt


qplot(x = friend_count, data = pf, binwidth = 0.4,
      color = I('black'),
       fill = I('#F79420')) + 
  scale_x_log10()

qplot(x = friend_count, data = pf, binwidth = 1,  
      color = I('black'), 
       fill = I('#F79420')) + 
  scale_x_sqrt(lim = c(0,1000))

p1 <- qplot(x = friend_count, data = pf) 
p2 <- qplot(x = log10(friend_count + 1), data = pf) 
p3 <- qplot(x = sqrt(friend_count), data = pf)

grid.arrange(p1,p2,p3,ncol=1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      binwidth = 10, geom = 'freqpoly', color = gender) + 
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50))
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x = www_likes, data =  subset(pf, !is.na(www_likes)), 
      geom = 'freqpoly', color = gender) + 
  scale_x_continuous() + 
  scale_x_log10()
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), 
       geom = 'boxplot') + 
  scale_y_continuous(lim = c(0,1000))

qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), 
       geom = 'boxplot') + 
  coord_cartesian(ylim = c(0,1000))
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), 
       geom = 'boxplot') + 
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y =friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') + 
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
names(pf)
qplot( x= gender, y = friendships_initiated, 
       data = subset(pf, !is.na(gender)), 
       geom = 'boxplot') + 
  coord_cartesian(ylim = c(0,125))

by(pf$friendships_initiated, pf$gender, summary)
```


Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0 )

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
length(pf$mobile_check_in)
sum(pf$mobile_check_in == 1)/ length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!