---
title: "problem_set_one_variable"
output: html_document
---

## Problem set one

```{r Diamonds}
library(ggplot2) # load ggplot
library(gridExtra)
diamonds <- data(diamonds) # load diamond set
names(diamonds) #list of variable names
summary(diamonds)
ls.str(diamonds)
```

## Including Plots

You can also embed plots, for example:

```{r histogram price}
qplot(x = price, data = diamonds, 
      color = I('black'), fill = I('#099DD9'))  #histogram
summary(diamonds$price) #summary of price
```


## Count by inequalities

Diamond Counts: 
Discover how to count by inequalities 
```{r diamonds counts}
names(diamonds)
sum(diamonds$price < 500) # number of things less 500 -- 1729
sum(diamonds$price < 250) # ` ` ` ` 250 --- 0 
sum(diamonds$price >= 15000) # ` ` ` ` 15000 1656
```

## Cheaper Diamonds
```{r cheaper diamonds}
qplot(x = price, data = diamonds, bindwidth = 1,
      color = I('black'), fill = I('#099DD9')) + 
  scale_x_continuous(breaks = seq(50, 1000, 50), lim = c(0, 1000))
```

# Price cut histogram 
```{r price cut histogram}
qplot(x = price, data = diamonds, bandwidth = 1,
      color = I('black'), fill = I('#099DD9')) + 
  facet_wrap(~cut,  ncol = 2) 
  # scale_x_log10()# this creates a column of factors based on cut and price
by(diamonds$price, diamonds$cut, summary, digits = 10) # need to do rounding here to see exact measurments. 
```

# Scales and multiple histograms
In the scales "free_y" or "free_x" let's each histogram 
display it's unique x and y coordinates instead of just 
one fixed position. 
```{r scales and multiple histograms}
qplot(x = price, data = diamonds,
  color = I('black'), fill = I('skyblue2')) + 
  facet_wrap(~cut, scales = 'free_y',  ncol =2)
```


#Price per carat by cut 
I wanted the binsize to 1/10 so i can get a nice distrobution of how the price*carat hold up.
```{r price per carat by cut}
qplot(x = price/carat, data = subset(diamonds, !is.na(price)), binwidth = .06,
      xlab = 'price per carat', 
      ylab = 'count of price per carat',
      color = I('black'), fill = I('skyblue2'))  + 
  facet_wrap(~cut, scales = 'free_y',  ncol = 2) + scale_x_log10()
```


#Price box plots 
Investigate the price of diamonds using box plots, 
numerical summaries, and one of the following categorical 
variables: cut, clarity or color. 
```{r box plots}
qplot(x = color, y = price, data = subset(diamonds, !is.na(price)), 
      xlab = 'color', ylab = 'price', 
      color = I('black'), fill = I('red'), geom = 'boxplot') + 
  coord_cartesian()

IQR(subset(diamonds, color == 'D')$price) #best color
IQR(subset(diamonds, color == 'J')$price) #worst color 
by(diamonds$price, diamonds$color, summary, digits = 10) # summary of quartiles
```

#Carat Frequency Polygon 
```{r Frequency Polygons}
qplot(x = carat,
      binwidth = 1,
      xlab = "carat", ylab = 'count',
      color = I('skyblue2'), fill = I('skyblue2'),
      data = diamonds, geom = 'freqpoly') +  
      scale_x_continuous() 

table(diamonds$carat) > 2000 # carat greater then count 2000 
?table
install.packages('tidyr')
install.packages('dplyr')
```


#gap minder data

Var: "state.abb"      "state.area"     "state.region"   "population"     "income"         "illiteracy"     "life.exp"       "murder"        
     "highSchoolGrad" "frost"          "area"          
```{r gapminder data}
library(reshape2)
require(gdata)
sdf <- read.csv('stateData.csv', header = T, 
                row.names = 1, check.names = F)
names(sdf) 
str(sdf)
qplot(x = income, data = sdf, 
      color = I('black'), fill = I('#099DD9')) 

qplot(x = life.exp, data = sdf, 
      color = I('black'), fill = I('#099DD9')) 

qplot(x = murder, data = sdf,
  color = I('black'), fill = I('skyblue2')) + 
  facet_wrap(~state.region, scales = 'free_y',  ncol =2)
```



# You friends birthdays


```{r birthdays}

```

