---
output:
  pdf_document: default
  html_document: default
---
Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)


qplot(x = age, y = friend_count, data = pf) #qplot function 
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() + xlim(13,90)
# xlim operation is a functional operation and you don't have to assign. 
# ggplot use aesthetic wrapper 
summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_jitter(alpha = 1/20) + xlim(13,90)
```

#### What do you notice in the plot?
Response: Essentially, what jitter does is sets a alpha parameter at .05 and shows us a more 
accurate uniform distrubtion of average friend counts across all ages. 

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13,90) + 
  coord_trans(y = "sqrt")
?coord_trans
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes: Essentially i took the base 10 log so  can see some more normal distrobution within the data set 
Alo I limited the y-axis by 1000 initiated friendship count so I could zoom in on the more denser regions. 
It seems like you get more initiated friends ships around the teenager to mid twenties mark and then it starts to go down.
This could be do to the fact that maybe peopel are more social active when they in this time frame (dates, new friendships, etc). 


```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/10, position = position_jitter(h = 0)) + 
  xlim(13,90) + 
  ylim(0, 1000) + 
  coord_trans(x = 'log10')
names(pf)
summary(pf$age)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)
age_groups <- group_by(pf, age) #this groups all data by age. 
pf.fc_by_age <- summarise(age_groups, #this creates a new dataframe with variabls as metrics. 
          friend_count_mean = mean(friend_count), 
          friend_count_median = median(friend_count), 
          n = n()) #n is the number of metrics in the age 

pf.fc_by_age <- arrange(pf.fc_by_age, age) #this arranges the age 

head(pf.fc_by_age,20)

'''pf.fc_by_age <- pf %.%  # using the %.% function will give us the same results as above so you do not have to call variables. 
  group_by(age) %.% 
  summarise(friend_count_mean = mean(friend_count), 
          friend_count_median = median(friend_count), 
          n = n()) %.%
  arrange(age)'''

```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line(color = "red", size = .5) + 
  coord_cartesian(xlim = c(13, 90)) 

?geom_line
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(aes(x=age, y=friendships_initiated), data=pf)+
geom_point(alpha=0.05, color='orange')+
xlim(13,90)+
coord_trans(y ='sqrt')+
geom_line(stat='summary', fun.y=mean)+
geom_line(stat='summary', fun.y=quantile, fun.args = list(probs = .1),
linetype=2, color='red') +
geom_line(stat = "summary", fun.y=quantile, fun.args = list(probs = .5), 
          color = "blue") +
geom_line(stat='summary', fun.y=quantile, fun.args = list(probs = .9),
linetype=2, color='darkgreen') + 
coord_cartesian(ylim = c(0,1000)) 
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:
Histograms show how much the over or under audience size.
***

### Correlation
Notes:

```{r Correlation}
cor(pf$age,pf$friend_count) #compute the correlation coeffience. 
cor.test(pf$age,pf$friend_count, method = "pearson") #this is better. 
with(pf, cor.test(age, friend_count, method = "pearson"))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count)) #subset the command of the data frame.
```

***

### Correlation Methods
Notes: The correlation methord (Pearson) measures the strenght between any two variables (x,y). 


***

## Create Scatterplots
Notes:

```{r}
names(pf)
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_line(alpha = 1/10, position = position_jitter(h = 0), color = I('red')) 

ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point() # funky outlies 
  
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  ylim(0, quantile(pf$likes_received, 0.95)) + 
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received, method = "pearson")) #correlation coef. 
```

Response:

***

### Moira on Correlation
Notes: Alot of data is correlated. Looking at FB data how many status updated someone posted in a month 
correlated with the last month the log in. They usually all kind of measure the same thing. Typically 
when she's workong on problem she wants to throw someone of this regression. 

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
names(Mitchell)
ggplot(aes(x = Temp, y = Month), data = Mitchell) + geom_point()
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Temp, y = Month), data = Mitchell) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
Zero correlation in this plot
b. What is the actual correlation of the two variables?
(Round to the thousandths place)

0.057

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Temp, Month, method = "pearson")) 
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point() + scale_x_continuous(breaks = seq(0,203, 12)) # break it into a sequence and step every 12 months
?scale_x_continuous

ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point() # this overlayes by 12 onths

x <- seq(0, 4*pi, pi/20)
y <- cos(x) 
qplot(x = x, y = y)
dcor.ttest(x,y)
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19,]
pf$age_with_months <- pf$age + (1 - pf$dob_month / 12) 
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))
pf
```

***

### Age with Months Means

```{r Age with Months Means}
names(pf)
library(dplyr)
age_with_months_groups <- group_by(pf,age_with_months)
pf.fc_by_age_months <- summarise(age_with_months_groups,
                                 friend_count_mean = mean(friend_count), 
                                 friend_count_median = median(friend_count), 
                                 n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months)
head(pf.fc_by_age_months) #create a age by months data months
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
# 852 months
pf.fc_by_age_months
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
  geom_line(color = 'red', sized = .5)

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age < 71)) + 
  geom_line() + 
  geom_smooth()

p2 <-ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count), 
             data = subset(pf, age < 71)) + 
  geom_line(stat = "summary", fun.y = mean)

library(gridExtra)
grid.arrange(p2,p1, p3,ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

