---
title: "R Notebook"
output: html_notebook
---

```{r price histograms with facet wrap}
library(ggplot2)
data(diamonds)
names(diamonds)
ggplot(data = diamonds, aes(x = log(price))) + 
  geom_histogram(aes(fill = cut)) + 
  facet_wrap(~color) + 
  scale_fill_brewer(palette = "Spectral")

```


thoughts: 
for the range, typically you are going to use the by function and look 
at the first and third quartile. 
```{r scatter plot}
ggplot(aes(x = table, y = price), data = diamonds) + 
  geom_point(aes(color = cut))  + 
  scale_fill_brewer(type = "qual")

by(diamonds$table, diamonds$cut, summary)
```


thoughts: 
what I want to do is create a variable in the data set 
called volume, which is the volume of the cuts. Secondly, 
I want to remove the top 1% of the diamonds so creat a subset and 
quantile. With this use a conditional statement. 
```{r Price vs Volume Scatter}
diamonds$volume <- (diamonds$x * diamonds$y * diamonds$z)
diamonds_volume_percent <- subset(diamonds, volume <= quantile(volume, .99)) #this gives me the lower 99% 
ggplot(aes(x = volume, y = price), data = subset(diamonds_volume_percent, volume != 0)) + 
  geom_point(aes(color = clarity)) + 
  scale_y_log10() + 
  scale_color_brewer(type = "div") + 
  xlim(0,400)



?quantile
```


Objective: 
# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

# This programming assignment WILL BE automatically graded.

```{r Explore many variables}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
pf$prop_initiated <- pf$friendships_initiated/pf$friend_count
qplot(x = friendships_initiated, y = prop_initiated , data = pf) + 
  geom_smooth()
```



```{r prop_initiaed vs tenure}
pf$year_joined <- floor(2014 - pf$tenure/365)
pf$year_joined.bucket <- cut(pf$year_joined, c(2004,2009,2011,2012,2014))
ggplot(aes(x = tenure, y = prop_initiated), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = median) + geom_smooth()

ggplot(aes(x = tenure, y = prop_initiated), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) + geom_smooth()


group1 <- subset(pf, year_joined.bucket == '(2012, 2014]')


```

# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ===========================================

```{r Price/Carat Binned, Facete}
ggplot(aes(x = cut, y = price/carat), data = diamonds) + 
  geom_point(aes(color = color)) + 
  facet_wrap(~clarity) + 
  scale_color_brewer(type = "div")
```

